---
title: "README"
author: "Jerad Acosta"
date: "October 26, 2014"
output: html_document
---

First we start by loading the data sets from the working directory
```{r}
## Load Data
# load test data
testdata <- read.table("./UCI HAR Dataset/test/X_test.txt")
label.index <- read.table("./UCI HAR Dataset/test/y_test.txt")
test.subjects <- read.table("./UCI HAR Dataset/test/subject_test.txt")
# load training data
traindata <- read.table("./UCI HAR Dataset/train/X_train.txt")
train.index <- read.table("./UCI HAR Dataset/train/y_train.txt")
train.subjects <- read.table("./UCI HAR Dataset/train/subject_train.txt")
# load features
features <- read.table("./UCI HAR Dataset/features.txt")
```

This code finds the appropriate Mean and Standard Deviation measurements
And then subsets the data to contain the observations of interest
Note the reasoning for how the subsetting was done.
For instance FreqMean was not considered since, according to the code book,
it contained the mean of frequencies in recording the data
as opposed to the mean of the actuals observations which we were interested in.
```{r}
# GOAL: find all mean calculations
# only using lower case mean to avoid angle measurements with 'Mean'
mean <- grepl("mean", features$V2)
# remove meanfreq since not mean of measurement but of frequency components
# as described in features_info.txt document
freq <- grepl("Freq", features$V2)

# subset features without meanfreq
features <- features[!(freq),]
#subset 46 features with mean
mean.features <- features[mean,]

## Goal: find all standard deviation calculations
std <- grepl("std", features$V2)
# subset 33 features calculating standard deviation
std.features <- features[std,]

# combine standard deviation and mean calculations
SnM.features <- rbind(mean.features, std.features)

# Subset data to desired measurements
measured.date <- traindata[,SnM.features$V1]
measured.date <- testdata[,SnM.features$V1]
```

Here we are transforming the indexes from the code book into
understandable classifiers for the label variable
```{r}
# convert label index to label names
test.label <- lapply(label.index,function(x) gsub(1,"Walking",x))
test.label <- lapply(test.label,function(x) gsub(2,"Walking_Upstairs",x))
test.label <- lapply(test.label,function(x) gsub(3,"Walking_Downstairs",x))
test.label <- lapply(test.label,function(x) gsub(4,"Sitting",x))
test.label <- lapply(test.label,function(x) gsub(5,"Standing",x))
test.label <- lapply(test.label,function(x) gsub(6,"Laying",x))

# bind labels to appropriate test data
testdata <- cbind(test.label, test.subjects, testdata)

# convert label index to label names
train.label <- lapply(train.index,function(x) gsub(1,"Walking",x))
train.label <- lapply(train.label,function(x) gsub(2,"Walking_Upstairs",x))
train.label <- lapply(train.label,function(x) gsub(3,"Walking_Downstairs",x))
train.label <- lapply(train.label,function(x) gsub(4,"Sitting",x))
train.label <- lapply(train.label,function(x) gsub(5,"Standing",x))
train.label <- lapply(train.label,function(x) gsub(6,"Laying",x))

traindata <- cbind(train.label, train.subjects, traindata)

# bind test data with training data
totaldata <- rbind(testdata, traindata)
```

Naming the Variables In accordance with tidy data
```{r}
# name variables
names(totaldata) <- c("Label", "Subject", as.character(SnM.features$V2))
```

This Creates new feature and labal text files for the new code book
```{r}
# create new feature text file for code book
write.table(SnM.features, file = "NewFeatures.txt")
write.table(totaldata$Label, file = "NewLabel.txt")
```
